Executive Constraints and Civil Conflict Onset

### 2020-02-02 paper.Rmd
First draft paper

[//]: # Setup R environment for entire doc
```{r include = F, results = "hide"}
library(corrplot)
library(dplyr)
library(extraDistr)
library(ggplot2)
library(gridExtra)
library(magrittr)
library(kableExtra)
library(knitr)
library(parallel)
library(precrec)
library(tidyr)

@@

options(knitr.kable.NA = ".")
opts_chunk$set(echo = F, eval = T, warning = F, fig.width = 5,
               fig.height = 3)

@@

@@

There are three general pathways which can theoretically link the
concept of executive constraints to the onset of civil conflict:
executive looting and state weakness, credible commitments to
bargained solutions, and the impact of violent repression by the state
on rebel mobilization. In the case of the first, the absence of
effective constraints enables leaders to loot state resources
[@Blattman2017], diverting funds for the purposes of personal
rent-seeking and corruption, which can undermine the administrative
and military capacity of the state to combat a nascent
insurgency. Furthermore, such environments allow leaders to pay
selective benefits to close supporters at the expense of wider
redistributive policies. To the extent that the provisioning of public
goods may ameliorate grievances against the state and co-opt potential
opposition, a reduced risk for civil conflict could be expected where
there is a high level of redistributive public spending and the
presence of quality institutions able to check the ability of the
executive to engage in cronyism [@Fjelde2009]. Finally, the lack of
accountability stemming from missing executive constraints could make
the state an attractive prize to predatory rebel groups; however,
prior empirical evidence lends little credence to this mechanism
[@Bazzi2014].

@@

The final pathway involves the impact of executive constraints on
shaping the ability of rebel groups to recruit new members. In order
to mount an effective challenge to the state, rebel leaders need to be
able to mobilize fighters while avoiding defections. However,
participating in an armed rebellion is a costly endeavour. Non-elites
face both the immediate risk of fighting as well as the possibility of
future reprisals by the government. Furthermore, given that most civil
conflicts are protracted events against an often times much stronger
opponent the short-term utility from participation is relatively
low. Thus, rebel leaders must be able to offer proportionate selective
benefits to offset the associated costs of fighting
[@Mason1996]. [^rational] These benefits can include not only economic
payoffs, but also protection from government violence. When the
government engages in violent repression, especially in the form of
indiscriminate violence, the cost of non-participation is
raised. Rebel groups can provide local security guarantees against
government reprisals such that the opportunity cost of fighting is
lowered, thereby incentivizing participation [@Regan2005]. However,
this is predicated on the government lacking the capacity to fully
eradicate any rebel group. Where that capacity exists the state is
free to engage in reprisals.

[^rational]: TODO: Assuming rational actors...

@@

Taken together, none of the aforementioned mechanisms are mutually
exclusive, and it is beyond the scope of the present paper to properly
adjudicate between them; however, they do converge on a single
testable proposition. Namely that higher levels of executive
constraints should be associated with a lowered risk of civil conflict
onset.  As discussed, this may be conditional though on the relative
capacity of the state --- if executive constraints affect the
mobilization potential for rebel groups, then in situations where the
government is willing to engage in repression and has the strength to
fully suppress the opposition, the impact of executive constraints on
civil conflict onset should be diminished. Furthermore, even if the
mobilization mechanism is not valid, the relationship could still be
confounded by the overall opportunity for rebellion --- the degree of
executive constraints may not be particularly relevant in relatively
stable countries with, for example, high levels of economic
development, little to no ethnic fractionalisation, and no easily
lootable resources to finance a rebellion. Regardless of
conditionalities, contrary results would indicate either that
executive constraints are indeterminate to the risk of civil conflict
or possibly that there is even a positive relationship since
constrained executives may be less able to effectively deal with
violent challenges due to risk aversion to violence and/or from the
slowdown of having to seek approval from a plurality of stakeholders
before committing to decisive action.

@@

ggplot(df, aes(year, value, colour = variable)) +
    geom_line() +
    theme(legend.title = element_blank()) +
    ylab("Percentage of Countries") +
    xlab("Year")

@@

As mentioned during the theoretical discussion of possible mechanisms,
executive constraints alone are insufficient in determining the
outbreak of conflict. One possible mediator is the relative capacity
of the state. For the present analysis I follow @Fearon2003 in using
GDP per capita as a proxy for the state's administrative and military
capacity by roughly capturing the revenue potential available to the
state. This is far from being an ideal proxy for state capacity; GDP
per capita could also capture economic grievances as well as the
relative ease of financing a rebellion given low wages
[@Hendrix2010]. Data on GDP per capita, however, is significantly more
complete than other common competing measures such as bureaucratic
quality or tax revenue as a percent of GDP.[^wbgi] Furthermore, it is
more correlated with the administrative aspects of state capacity than
measures focused solely on military strength, such as military
expenditures [@Hendrix2010]. Thus, I use the natural log of the
*RGDP^o* variable, constant price real GDP, from Penn World Tables
(PWT) v9.1 (1950-2017) to allow for both cross-country and
cross-temporal comparisons of GDP per capita [@Feenstra2015] and
include an interaction term between GDP per capita and the latent
measure for executive constraints to capture the conditional
relationship.

[^wbgi]: For example, the World Bank governance indicator on
    government effectiveness extends back only until 1996 (TODO:
    cite). Given that measures of institutional quality often change
    fairly slowly, punctuated only by abrupt events such as coups or
    revolutions, an extended time series is desirable. Furthermore, in
    the specific case of bureaucratic quality, there is the added
    issue of endogeneity. An unconstrained executive may seek to
    promote government corruption and undermine both meritocratic
    recruitment into the bureaucracy as well as its impartiality.

In addition to state capacity, I also include a number of standard
variables drawn largely from previous literature. The list of
variables are not meant to provide an all encompassing theory of civil
conflict, but rather roughly capture the opportunity structures for
armed conflict.  This includes a single year lagged measure of GDP per
capita growth from PWT in order to capture short term economic shocks,
which may affect both the strength and legitimacy of the state
[@Bazzi2014]. I also include a measure of mean elevation from the
GROWup dataset [@Girardin2015] to capture terrain advantageous for
conducting an insurgency [@Fearon2003] as well as a measure of
population density derived from population data from PWT and the area
in square kilometers for each country from GROWup. Countries with a
higher population density may require a more extensive and competent
central administration and policing force in order to suppress violent
dissent [@Fearon2003]. There's also the possibility that a high
population density increases the competition for resources. If the
state is identified as the source of a particular scarcity, the risk
of civil conflict may increase. Similarly to GDP per capita, I also
transform mean elevation and population by the natural log. Finally, I
include measures of ethnic fractionalization in the form of the number
of political relevant ethnic groups and the percent of the population
facing discrimination. Where there are significant ethnic cleavages
and state sponsored discrimination, ethnic ties can facilitate
collective action thereby lowering the mobilization costs of rebel
groups [@Cederman2011].

@@

The analysis of the results is split into several sections. First, the
parameter estimates from the executive constraints measurement model
will be presented followed by the parameters pertaining to the
logistic regression and an assessment of model fit. The final section
will be a discussion linking the results back to the outlined theories
as well as an overview of the limitations of the presented model.

@@

```{r  fig.cap = "\\label{fig:fpars} hello", warning = F}
load("./posteriors/full_model/input.RData")
fa <- read_post("./posteriors/full_model/fa_posteriors.csv.gz")

@@

p1 <- plot_pars(lambda)
p2 <- plot_pars(psi)
p3 <- plot_pars(gamma)
p4 <- plot_pars(delta_eta)

@@

```{r fa, fig.height = 6, fig.width = 6, warning = F}
fa_pars <- extract(fa, c("lambda", "gamma", "psi", "delta", "eta")) %>%
    post_summarise(probs = c(0.10, 0.9)) %>%
    mutate(parameter = sub("\\[\\d+\\]", "", parameter) %>% sprintf("$\\%s$", .),
           interval = sprintf("[%.3f, %.3f]", signif(`10%`, 3), signif(`90%`, 3)),
           manifest = c(rep(c(lg, nonlg), 3),
                        rep("v2lgbicam", 2))) %>%
    select(-`10%`, -`90%`) %>%
    pivot_wider(id_cols = manifest, names_from = parameter, values_from = interval)

@@

```{r fig.width = 3, fig.height = 2, fig.cap = "\\label{fig:bicam} Posterior density plots for $\\theta$"}
theta <- extract(fa, "theta")
plot_dens(theta, groups = lgbicam)

@@

```{r fig.cap = "\\label{fig:fa-line} Error"}
err <- read_post("posteriors/full_model/err_posteriors.csv.gz")

nonlg_est <- extract(err, "nonlg_est") %>%
    post_summarise %>%
    rename(codelow = `2.5%`, median = `50%`, codehigh = `97.5%`) %>%
    mutate(interval = codehigh - codelow,
           se = nonlg_mm$se[, 1],
           obs = nonlg_mm$obs[, 1],
           diff = obs - median)

@@

x <- data.frame(point = c(nonlg_est$obs[idx], nonlg_est$median[idx]),
                group = rep(1:length(idx), 2),
                factor = c(rep(1, length(idx)), rep(2, length(idx))) %>% as.factor,
                theta = rep(theta_medians, 2))

@@

ggplot(x, aes(theta, point, group = group, col = factor)) +
    geom_line(alpha = 0.3) +
    geom_point(alpha = 0.8) +
    geom_line(data = y, aes(theta, yhat, group = NA, col = NA), alpha = 1)

@@

```{r fig.cap = "\\label{fig:err} Errors"}
p1 <- ggplot(nonlg_est, aes(se, interval)) +
    geom_point(col = "#56B4E9", alpha = 0.2, shape = 1)

p2 <- ggplot(nonlg_est, aes(se, diff)) +
    geom_point(col = "#56B4E9", alpha = 0.2)

@@

```{r fig.width = 7, fig.height = 5, fig.cap = "\\label{fig:theta} Theta"}
theta_summarised <- post_summarise(theta, probs = c(0.10, 0.5, 0.90)) %>%
    rename(codelow = `10%`, median = `50%`, codehigh = `90%`)

@@

filter(theta_summarised, country_name %in% targets) %>%
    ggplot(aes(year, median, fill = country_name)) +
    geom_line() +
    geom_ribbon(aes(ymin = codelow, ymax = codehigh), alpha = 0.2)

@@

```{r fig.cap = "\\label{fig:tie} Tie Fighter"}
targets <- c("Algeria", "Argentina", "Armenia", "Chile", "Colombia",
             "Egypt", "Georgia", "Ghana", "Hungary", "Indonesia",
             "Iran", "Israel", "Japan", "Kenya", "Mexico", "Poland",
             "Portugal", "Romania", "Rwanda", "Turkey", "Tunisia",
             "United States of America", "Zimbabwe")

df <- filter(theta_summarised, year %in% c(2017, 2007),
             country_name %in% targets) %>%
    mutate(year = as.factor(year),
           country_name = ifelse(country_name == "United States of America",
                                 "USA", country_name) %>%
               factor(levels = sort(c(targets, "USA"), T)))

ggplot(df, aes(median, country_name, colour = year)) +
    geom_point() +
    geom_errorbarh(aes(xmax = codehigh, xmin = codelow))

@@

```{r results = "hide"}
romania_idx <- filter(final.df, country_name == "Romania", year %in% c(2007, 2017)) %$%
    reduced_idx

romania_p <- signif(mean(theta[, romania_idx[1]] > theta[, romania_idx[2]]) * 100, 2)

chile_idx <- filter(final.df, country_name == "Chile", year %in% c(2007, 2017)) %$%
    reduced_idx

chile_p <- signif(mean(theta[, chile_idx[1]] < theta[, chile_idx[2]]) * 100, 2)
```

To provide a further snapshot of a greater number of countries, figure
\ref{fig:tie} plots the $\theta$ estimates for 2017, the last year in
the dataset, compared to 2007 by country. A full list of all countries
is available in [Appendix C](#appendix-c). Note, the statistical
significance of differences between years for a single country cannot
strictly be judged by overlapping credible intervals
[@Schenker2001]. For example, Romania, despite having overlapping
credible intervals, still has a posterior probability of approximately
$`r romania_p`\%$ that the estimate for 2017 is lower than 2007. In
the case of Chile with even more visual overlap, there is a
$`r chile_p`\%$ posterior probability that the estimate for 2017 is
higher than 2007.

@@

```{r reg, fig.height = 3, warning = F}
reg <- read_post("posteriors/full_model/reg_posteriors.csv.gz")

@@

```{r fig.cap = "\\label{fig:reg} hello world"}
plot_pars(beta)
```

@@

```{r results = "hide"}
merged.df <- readRDS("data/merged_data.rds")

gdp_nigeria <- filter(merged.df, country_name == "Nigeria", year == 2014) %$%
    rgdpepc %>%
    prettyNum(big.mark = ",")

@@

```{r results = "hide"}
gdp_eth <- filter(merged.df, country_name == "Ethiopia", year == 2012) %$%
    rgdpepc %>%
    prettyNum(big.mark = ",")

gdp_india <- filter(merged.df, country_name == "India", year == 2014) %$%
    rgdpepc %>%
    prettyNum(big.mark = ",")

gdp_nz <- filter(merged.df, country_name == "New Zealand", year == 2002) %$%
    rgdpepc %>%
    prettyNum(big.mark = ",")
```

To interpret the results for the interaction term between the slopes
for executive constraints and GDP per Capita I calculate the marginal
effect of $\theta$, executive constraints, on the predicted
probability of civil conflict onset for different levels of GDP per
capita while holding constant the other covariates.  In this case, I
took the mean of GDP per capita growth, population density, mean
elevation, politically relevant ethnic group count, and peace years
while setting the dummy indicators for independence, ongoing conflict,
and neighbouring conflict and the random intercepts for year and
country to 0. The levels of GDP per capita were selected based on the
10\%, 50\%, and 90\% quantiles of the standardized log transformed
values. These levels roughly correspond to the GDP per capita of
Ethiopia in 2012 at $`r gdp_eth`$, India in 2014 at $`r gdp_india`$,
and New Zealand in 2002 at $`r gdp_nz`$. In the dataset there are $75$
conflict episode onsets below the first level, $198$ between the first
and second, $73$ between the second and third, and $6$ above the
third.

@@

state <- quantile(final.df$rgdpepc, probs = c(0.1, .5, 0.9))

f <- function(x, y) {
    with(final.df, alpha +
                   beta[, 1] * x +
                   beta[, 2] * y +
                   beta[, 3] * mean(rgdpepc_gro) +
                   beta[, 4] * mean(pop_density) +
                   beta[, 5] * mean(meanelev) +
                   beta[, 6] * mean(rlvt_groups_count) +
                   beta[, 7] * 0 +
                   beta[, 8] * 0 +
                   beta[, 9] * mean(peace_yrs) +
                   beta[, 10] * 0 +
                   beta[, 11] * x * y)
}

df <- lapply(state, function(y) {
    mclapply(theta_seq, function(theta) {
        p_hat <- boot::inv.logit(f(theta, y)) * 100
        post_summarise(p_hat, probs = c(0.1, 0.5, 0.9)) %>%
            mutate(theta = theta,
                   y = y)
    }, mc.cores = 4) %>% bind_rows
}) %>% bind_rows

@@

ggplot(df, aes(theta, median)) +
    geom_line() +
    geom_ribbon(aes(ymin = codelow, ymax = codehigh), alpha = 0.2) +
    facet_grid(. ~ y) +
    xlab("Executive Constraints") +
    ylab("Predicted Probability")
```

@@

plot_dens(p_hat, groups = final.df$lepisode_onset) + xlim(c(0, 0.4))
```

```{r psum, warning = F}
p_mean <- colMeans(p_hat)

@@

```{r fig.width = 7, fig.cap = "\\label{fig:roc} hello"}

@@

### Discussion

- Link back to original theory, confirmatory

- Small effect size
- Poor fit
- Noisy Data
- Robustness?
- Compared to democracy?

Further analyses?

# Conclusion {-}

>>>>

[//]: # Setup R environment for entire doc
```{r include = F, results = "hide"}
library(corrplot)
library(dplyr)
library(extraDistr)
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(magrittr)
library(kableExtra)
library(knitr)
library(parallel)
library(precrec)
library(tidyr)

@@

options(knitr.kable.NA = ".")
opts_chunk$set(echo = F, eval = T, warning = F, fig.width = 5,
               fig.height = 3)
```

@@

Since the end of World War II, civil conflict has been the deadliest
and most prevalent form of conflict. In that time period, upwards to
two-thirds of all countries have experienced at least one minor
insurgency and approximately one-third of all countries have
experienced a major civil war with at least 1,000 battle deaths
[@Pettersson2019]. Given the impact of civil conflict and the
increasing availability of conflict related data, the last several
decades has witnessed a veritable explosion in the number of
quantitative studies investigating the causes of civil conflict
[@Blattman2017]. Scholars have proposed numerous theories ranging
from motivation-based arguments --- such as, for example, the role of
horizontal inequalities [@Cederman2011] --- to opportunity-based
explanations centered on economic factors [@Collier2012; @Collier2009]
as well as the political and structural conditions underlying state
weakness [@Hegre2001; @Fearon2003]. This latter strain of research in
particular has spurred extensive work into examining the
distinguishing features of weak states and whether political
institutions play a significant role in the risk of civil conflict.
Several studies in this vein have argued in favor of institutions
utilizing disaggregated measures to also investigate the specific
types of institutions relevant to the study of civil conflict
[@Taydas2010; @Hegre2015a; @Bartusevic2018].

This paper extends that work by taking a disaggregated approach to
studying the effect of executive constraints on the risk of civil
conflict onset. The motivation to specifically focus on executive
constraints is two-fold: first, the executive is often times one of
the key decision-makers in a state's regime. Institutions that provide
a check against the types of actions which may exacerbate the risk of
rebellion should be relevant to the study of civil conflict. In doing
we can further our understanding of how political institutions shape
the incentives for violence, especially through the influence of
institutions on state behaviour.  Second, the link between executive
constraints and civil conflict is relatively understudied. The
previous literature that exists is fairly limited and contradictory in
terms of results [@Getmansky2013; @Fjelde2009].

The principle finding is that there is tentative evidence that
executive constraints do matter and that there is a negative
relationship between the latent level of constraints on the executive
and the risk of civil conflict onset. However, this relationship
largely disappears for low-income country-years when an interaction
term with GDP per capita is included. If the primary theoretical
mechanism involves the impact of executive constraints on rebel
mobilization --- an unchecked executive may be more likely to engage
in indiscriminate violence which could lead to a loss of legitimacy as
well as the ability of rebel groups to offer local security guarantees
to attract recruits --- then the incentives for violence in low-income
country-years may be so great that it does not matter how constrained
the executive may or may not be.

We should not be overconfident with these results, however. Similar to
most other studies examining conflict onset, the model is a poor fit
to the underlying data. Modeling civil conflict is a difficult task,
especially when trying to capture the dynamics of conflict with a
standard time series cross-country regression. The results will
require verification by additional studies using alternative
methodological approaches.

In what follows, a deeper justification for why political institutions
are relevant to the study of civil conflict together with an overview
of related empirical literature will be presented. The theoretical
mechanisms linking executive constraints to conflict onset will then
be discussed followed by an empirical analysis for which a latent
estimate for executive constraints will be estimated and used in a
cross-country time series logistic regression predicting conflict
onset.

@@

There are three general pathways which can theoretically link the
level of executive constraints to the onset of civil conflict:
executive looting and state weakness, credible commitments to
bargained solutions, and the impact of violent repression by the state
on rebel mobilization. In the case of the first, the absence of
effective constraints enables leaders to loot state resources
[@Blattman2017], diverting funds for the purposes of personal
rent-seeking and corruption, which can undermine the administrative
and military capacity of the state to combat a nascent
insurgency. Furthermore, such environments allow leaders to pay
selective benefits to close supporters at the expense of wider
redistributive policies. To the extent that the provisioning of public
goods may ameliorate grievances against the state and co-opt potential
opposition, a reduced risk for civil conflict could be expected where
there is a high level of redistributive public spending and the
presence of quality institutions able to check the ability of the
executive to engage in cronyism [@Fjelde2009]. Finally, the lack of
accountability stemming from missing executive constraints could make
the state an attractive prize to predatory rebel groups; however,
prior empirical evidence lends little credence to this mechanism
[@Bazzi2014].

@@

The final pathway involves the impact of executive constraints on
shaping the ability of rebel groups to recruit new members. In order
to mount an effective challenge to the state, rebel leaders need to be
able to mobilize fighters while avoiding defections. However,
participating in an armed rebellion is a costly endeavour. Non-elites
face both the immediate risk of fighting as well as the possibility of
future reprisals by the government. Furthermore, given that most civil
conflicts are protracted events against an often times much stronger
opponent the short-term utility from participation is relatively
low. Thus, rebel leaders must be able to offer proportionate selective
benefits to offset the associated costs of fighting
[@Mason1996]. These benefits can include not only economic payoffs,
but also protection from government violence. When the government
engages in violent repression, especially in the form of
indiscriminate violence, the cost of non-participation is
raised. Rebel groups can provide local security guarantees against
government reprisals such that the opportunity cost of fighting is
lowered, thereby incentivizing participation [@Regan2005]. However,
this is predicated on the government lacking the capacity to fully
eradicate any rebel group. Where that capacity exists the state is
free to engage in reprisals.

@@

Taken together, none of the aforementioned mechanisms are mutually
exclusive, and it is beyond the scope of the present paper to properly
adjudicate between them; however, they do converge on a single
testable proposition. Namely that higher levels of executive
constraints should be associated with a lowered risk of civil conflict
onset. This proposition can be further extended if state weakness is a
primary driver of civil conflict and the relative capacity of the
state is effected by the levels of executive constraints, as depicted
in the first theoretical pathway regarding executive looting, then
state capacity could have a mediating affect on the relationship
between executive constraints and conflict onset. Another extension
would be if executive constraints affect the mobilization potential
for rebel groups, then there could also be a confounding effect from
the relative threshold for mobilization. Finally, the relationship
could still be confounded by the overall opportunity for rebellion ---
the degree of executive constraints may not be particularly relevant
in relatively stable countries lacking the conditions conducive to
carrying out a rebellion. Regardless of conditionalities, contrary
results would indicate either that executive constraints are
indeterminate to the risk of civil conflict or possibly that there is
even a positive relationship since constrained executives may be less
able to effectively deal with violent challenges due to risk aversion
to violence and/or from the slowdown of having to seek approval from a
plurality of stakeholders before committing to decisive action.

@@

ggplot(df, aes(year, value, colour = variable)) +
    geom_line() +
    theme_tufte() +
    scale_color_colorblind() +
    theme(legend.title = element_blank()) +
    ylab("Percentage of Countries") +
    xlab("Year")

@@

As mentioned during the theoretical discussion of possible mechanisms,
executive constraints alone are insufficient in determining the
outbreak of conflict. One possible mediator is the relative capacity
of the state. For the present analysis I follow @Bartusevic2018 in
using the Composite Indicator of National Capability (CINC) from the
National Material Capabilities v5.0 dataset (1816-2012) [@Singer1972;
@Singer1987]. The CINC index is aggregated from several demographic,
industrial, and military indicators in order to capture the available
resources a country can bring to bear in the event of a conflict. This
is an imperfect measure that largely misses the state's bureaucratic
and administrative capacity [@Hendrix2010]. [^wbgi] However, given
that the dimension of interest for this analysis is the repressive
capacity of the state, material capabilities, especially military
resources, should still provide an adequate proxy.

[^wbgi]: Furthermore, a common issue with many measures of
    bureaucratic quality is missingness.  For example, the World Bank
    governance indicator on government effectiveness extends back only
    until 1996 (TODO: cite). Given that measures of institutional
    quality often change fairly slowly, punctuated only by abrupt
    events such as coups or revolutions, an extended time series is
    desirable. Furthermore, in the specific case of bureaucratic
    quality, there is the added issue of endogeneity. An unconstrained
    executive may seek to promote government corruption and undermine
    both meritocratic recruitment into the bureaucracy as well as its
    impartiality.

To capture the relative ease of mobilizing fighters, especially in
terms of being able to offer economic selective benefits, I include
GDP per capita. Specifically, I use the natural log of the $RGDP^o$
variable, constant price real GDP per capita, from Penn World Tables
(PWT) v9.1 (1950-2017) to allow for both cross-country and
cross-temporal comparisons of GDP per capita [@Feenstra2015]. The
justification for taking the natural log is that the marginal effect
of the variable should decrease with higher values. In other words,
the difference between the effects of relatively high levels of GDP
per capita should be significantly smaller than the difference between
the effects of low levels of GDP per capita. Finally, I also include
an interaction term between GDP per capita and the latent measure for
executive constraints to capture the potential conditional
relationship.

In addition to the previous indicators, I also include a number of
standard covariates drawn largely from previous literature. The list
of indicators are not meant to provide an all encompassing theory of
civil conflict, but rather roughly capture the opportunity structures
for rebellion.  This includes a single year lagged measure of GDP per
capita growth from PWT in order to capture short term economic shocks,
which may affect both the strength and legitimacy of the state
[@Bazzi2014]. I also include a measure of mean elevation from the
GROWup dataset [@Girardin2015] to capture terrain advantageous for
conducting an insurgency [@Fearon2003] as well as a measure of
population density derived from population data from PWT and the area
in square kilometers for each country from GROWup. Countries with a
higher population density may require a more extensive and competent
central administration and police force in order to suppress violent
dissent [@Fearon2003]. There's also the possibility that a high
population density increases the competition for resources. If the
state is identified as the source of a particular scarcity, the risk
of civil conflict may increase. Similar to GDP per capita, I also
transform mean elevation and population by the natural log with the
expectation that the scale increase should diminish as values get
larger. Finally, I include measures of ethnic fractionalization in the
form of the number of political relevant ethnic groups and the percent
of the population facing discrimination. Where there are significant
ethnic cleavages and state sponsored discrimination, ethnic ties can
facilitate collective action thereby lowering the mobilization costs
of rebel groups [@Cederman2011].

@@

The analysis of the results is split into several sections. First, the
parameter estimates from the executive constraints measurement model
will be presented. This will then be followed by the parameters
pertaining to the logistic regression before concluding with an
assessment of model fit.

@@

```{r  fig.cap = "\\label{fig:fpars} Measurement Model Parameter Estimates"}
load("./posteriors/full_model/input.RData")
fa <- read_post("./posteriors/full_model/fa_posteriors.csv.gz")

@@

p1 <- plot_pars(lambda, "Lambda")
p2 <- plot_pars(psi, "Psi")
p3 <- plot_pars(gamma, "Gamma")
p4 <- plot_pars(delta_eta, "")

@@

```{r fa, fig.height = 6, fig.width = 6}
fa_pars <- extract(fa, c("lambda", "gamma", "psi", "delta", "eta")) %>%
    post_summarise(probs = c(0.10, 0.9)) %>%
    mutate(parameter = sub("\\[\\d+\\]", "", parameter) %>% sprintf("$\\%s$", .),
           interval = sprintf("[%.3f, %.3f]", signif(`10%`, 3), signif(`90%`, 3)),
           manifest = c(rep(c(lg, nonlg), 3),
                        rep("v2lgbicam", 2))) %>%
    select(-`10%`, -`90%`) %>%
    pivot_wider(id_cols = manifest, names_from = parameter, values_from = interval)

@@

```{r fig.width = 6, fig.height = 2, fig.cap = "\\label{fig:bicam} Posterior density plots for executive constraints"}
theta <- extract(fa, "theta")

groups <- ifelse(lgbicam == 1, "Legislature Present", "No Legislature")
plot_dens(theta, groups = groups)

@@

```{r fig.cap = "\\label{fig:fa-line} Posterior shrinkage for \\textit{v2juncind}"}
err <- read_post("posteriors/full_model/err_posteriors.csv.gz")

target <- colnames(nonlg_mm$obs)[1]
nonlg_est <- extract(err, "nonlg_est") %>%
    post_summarise %>%
    rename(codelow = `2.5%`, median = `50%`, codehigh = `97.5%`) %>%
    mutate(interval = codehigh - codelow,
           se = nonlg_mm$se[, 1],
           obs = nonlg_mm$obs[, 1],
           diff = obs - median)

@@

x <- data.frame(point = c(nonlg_est$obs[idx], nonlg_est$median[idx]),
                group = rep(1:length(idx), 2),
                factor = c(rep("Observed Point Estimate", length(idx)),
                           rep("Estimated Posterior Median", length(idx))) %>% as.factor,
                theta = rep(theta_medians, 2))

@@

ggplot(x, aes(theta, point, group = group, col = factor)) +
    geom_line(alpha = 0.3) +
    geom_point(alpha = 0.8) +
    geom_line(data = y, aes(theta, yhat, group = NA), colour = "black", alpha = 1) +
    theme_tufte() +
    scale_color_colorblind() +
    theme(legend.title = element_blank()) +
    ylab(target) +
    xlab("Theta")

@@

```{r fig.cap = "\\label{fig:err} Error plots"}
p1 <- ggplot(nonlg_est, aes(se, interval)) +
    geom_point(col = "#56B4E9", alpha = 0.2, shape = 1) +
    theme_tufte() +
    ylab("Posterior CI Range") +
    xlab("Observed Error")

p2 <- ggplot(nonlg_est, aes(se, diff)) +
    geom_point(col = "#56B4E9", alpha = 0.2) +
    theme_tufte() +
    ylab("Pointwise difference") +
    xlab("Observed Error")

@@

```{r fig.width = 7, fig.height = 5, fig.cap = "\\label{fig:theta} Executive Constraints Over Time"}
theta_summarised <- post_summarise(theta, probs = c(0.10, 0.5, 0.90)) %>%
    rename(codelow = `10%`, median = `50%`, codehigh = `90%`)

@@

filter(theta_summarised, country_name %in% targets) %>%
    ggplot(aes(year, median, fill = country_name)) +
    theme_tufte() +
    scale_fill_colorblind() +
    geom_line() +
    geom_ribbon(aes(ymin = codelow, ymax = codehigh), alpha = 0.2) +
    theme(legend.title = element_blank())

@@

```{r fig.cap = "\\label{fig:tie} Executive Constraints 2000 vs 2010"}
targets <- c("Algeria", "Argentina", "Armenia", "Chile", "Colombia",
             "Egypt", "Georgia", "Ghana", "Hungary", "Indonesia",
             "Iran", "Israel", "Japan", "Kenya", "Mexico", "Poland",
             "Portugal", "Romania", "Rwanda", "Turkey", "Tunisia",
             "United States of America", "Zimbabwe")

df <- filter(theta_summarised, year %in% c(2010, 2000),
             country_name %in% targets) %>%
    mutate(year = as.factor(year),
           country_name = ifelse(country_name == "United States of America",
                                 "USA", country_name) %>%
               factor(levels = sort(c(targets, "USA"), T)))

ggplot(df, aes(country_name, median, colour = year)) +
    geom_errorbar(aes(ymax = codehigh, ymin = codelow),
                  position = position_dodge(width = 0.5)) +
    theme_tufte() +
    scale_color_colorblind() +
    theme(axis.title.y = element_blank(),
          legend.title = element_blank()) +
    ylab("Theta") +
    coord_flip()

@@

```{r results = "hide"}
chile_idx <- filter(final.df, country_name == "Chile", year %in% c(2000, 2010)) %$%
    reduced_idx

chile_p <- signif(mean(theta[, chile_idx[1]] < theta[, chile_idx[2]]) * 100, 2)

tunisia_idx <- filter(final.df, country_name == "Tunisia", year %in% c(2000, 2010)) %$%
    reduced_idx

tunisia_p <- signif(mean(theta[, tunisia_idx[1]] < theta[, tunisia_idx[2]]) * 100, 2)
```

To provide a further snapshot of a greater number of countries, figure
\ref{fig:tie} plots the $\theta$ estimates for 2010, the last year in
the dataset, compared to 2000 by country. A full list of all countries
is available in [Appendix C](#appendix-c). Note, the statistical
significance of differences between years for a single country cannot
strictly be judged by overlapping credible intervals
[@Schenker2001]. For example, Chile, despite having overlapping
credible intervals, still has a posterior probability of approximately
$`r chile_p`\%$ that the estimate for 2010 is lower than 2000. In the
case of Tunisia with even more visual overlap, there is a
$`r tunisia_p`\%$ posterior probability that the estimate for 2010 is
higher than 2000.

@@

```{r reg, fig.height = 3}
reg <- read_post("posteriors/full_model/reg_posteriors.csv.gz")

@@

```{r fig.cap = "\\label{fig:reg} Regression parameters"}
plot_pars(beta, hline = T)
```

@@

```{r results = "hide"}
gdp_nigeria <- filter(merged.df, country_name == "Nigeria", year == 2014) %$%
    rgdpepc %>%
    prettyNum(big.mark = ",")

@@

```{r results = "hide"}
gdp_rwa <- filter(merged.df, country_name == "Rwanda", year == 2007) %$%
    rgdpepc %>%
    prettyNum(big.mark = ",")

gdp_phl <- filter(merged.df, country_name == "Philippines", year == 2008) %$%
    rgdpepc %>%
    prettyNum(big.mark = ",")

gdp_po <- filter(merged.df, country_name == "Portugal", year == 2009) %$%
    rgdpepc %>%
    prettyNum(big.mark = ",")
```

To interpret the results for the interaction term between the slopes
for executive constraints and GDP per Capita I calculate the marginal
effect of $\theta$, executive constraints, on the predicted
probability of civil conflict onset for different levels of GDP per
capita while holding constant the other covariates.  In this case, I
took the mean of GDP per capita growth, population density, mean
elevation, politically relevant ethnic group count, and peace years
while setting the dummy indicators for independence, ongoing conflict,
and neighbouring conflict and the random intercepts for year and
country to 0. The levels of GDP per capita were selected based on the
10\%, 50\%, and 90\% quantiles of the standardized log transformed
values. These levels roughly correspond to the GDP per capita of
Rwanda in 2007 at $`r gdp_rwa`$, the Philippines in 2008 at $`r gdp_phl`$,
and Portugal in 2009 at $`r gdp_po`$. In the dataset there are 60
conflict episode onsets below the first level, $152$ between the first
and second, $55$ between the second and third, and $4$ above the
third.

@@

state <- quantile(X[, "rgdpepc"], probs = c(0.1, .5, 0.9))

f <- function(x, y) {
    with(as.data.frame(X),
         alpha +
         beta[, 1] * x +
         beta[, 2] * y +
         beta[, 3] * mean(rgdpepc_gro) +
         beta[, 4] * mean(cinc) +
         beta[, 5] * mean(discrimpop) +
         beta[, 6] * mean(pop_density) +
         beta[, 7] * mean(meanelev) +
         beta[, 8] * mean(rlvt_groups_count) +
         beta[, 9] * 0 +
         beta[, 10] * 0 +
         beta[, 11] * mean(peace_yrs) +
         beta[, 12] * 0 +
         beta[, 13] * x * y)
}

df <- lapply(state, function(y) {
    mclapply(theta_seq, function(point) {
        p_hat <- boot::inv.logit(f(point, y)) * 100
        post_summarise(p_hat, probs = c(0.1, 0.5, 0.9)) %>%
            mutate(theta = point,
                   y = y)
    }, mc.cores = 4) %>% bind_rows
}) %>% bind_rows

@@

ggplot(df, aes(theta, median)) +
    geom_line() +
    geom_ribbon(aes(ymin = codelow, ymax = codehigh), alpha = 0.2) +
    facet_grid(. ~ y) +
    theme_tufte() +
    xlab("Executive Constraints") +
    ylab("Predicted Probability")

@@

groups <- ifelse(final.df$lepisode_onset == 1,
                 "Actual Onset",
                 "No Onset")
plot_dens(p_hat, groups = groups) +
    xlim(c(0, 0.4))
```

```{r psum}
p_mean <- colMeans(p_hat)

@@

```{r fig.width = 7, fig.cap = "\\label{fig:roc} ROC & Precision-Recall Plots"}

@@

# Discussion {-}

How do the results from the presented model relate back to the
theories potentially linking executive constraints to civil conflict?
According to the model there is support for the general proposition
that the risk of civil conflict onset decreases the more constrained
the executive is by other government branches when controlling for the
selected covariates. As stated previously, however, which theoretical
mechanism(s) might be responsible for this relationship lies outside
the scope of this paper in large part because the mechanisms that were
discussed are not mutually exclusive. Furthermore, quantitative
country-year regressions such as this one can only detect general
patterns and provide evidence towards the associations of particular
indicators. What they cannot do is examine the microfoundations
underpinning a particular theory.
